---
title: "MortalityGaps R Package"
author: "Marius Pascariu"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
  # rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette --- DoubleGap R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package contains source code for the Double-Gap model for forecasting 
life expectancy in human populations.


# Description
Life expectancy is highly correlated over time among countries and 
between males and females. These associations can be used to improve forecasts. 
Here we have implemented a method for forecasting female life expectancy based on 
analysis of the gap between female life expectancy in a country compared with
the record level of female life expectancy in the world. Second, to forecast 
male life expectancy, the gap between male life expectancy and female life 
expectancy in a country is analysed. We named this method the Double-Gap model.
For a detailed description of the method see Pascariu et al. (2017).

## Installation

1. Make sure you have the most recent version of R
2. Run the following code in your R console 

```R
install.packages("MortalityGaps")
```

## Updating to the latest version of the package

You can track and contribute to the development of `MortalityGaps` on [GitHub](https://github.com/mpascariu/MortalityLaws). To install it:

1. Install the release version of `devtools` from CRAN with `install.packages("devtools")`.

2. Make sure you have a working development environment.
    * **Windows**: Install [Rtools](https://CRAN.R-project.org/bin/windows/Rtools/).
    * **Mac**: Install `Xcode` from the Mac App Store.
    * **Linux**: Install a compiler and various development libraries (details vary across different flavors of Linux).

3. Install the development version of `MortalityGaps`.

   ```R
   devtools::install_github("mpascariu/MortalityGaps")
   ```

# Help
All functions are documented in the standard way, which means that 
once you load the package using ```library(MortalityGaps)```
you can just type ```?DoubleGap``` to see the help file. 


# Examples

```{r}
library(MortalityGaps)
```

### Input data 
```{r}
# Collection of life expectancies for female populations
exF <- MortalityGaps.data$exF
# Life expectancy for male populations
exM <- MortalityGaps.data$exM

head(exF)
```


### Fit DG model at age 0 for Australia using data from 1950 to 2014
```{r}
M0 <- DoubleGap(DF = exF,
                DM = exM,
                age = 0,
                country = "AUS",
                years = 1950:2014)
M0
```

### Summary results
```{r}
summary(M0)
```

### Forecast life expectancy in Australia until 2050
```{r}
P0 <- predict(M0, h = 36)
```

### Plot the results
```{r, fig.asp = 0.9}
plot(P0)
```





